---
title: "Google Trends and Births"
author: "Pekka Haimi"
date: "2022-11-10"
output:
  html_document:
    keep_md: yes
---
This is analysis part of Master Thesis written by Pekka Haimi.  
Aim of the analysis:

* Import Finnish birth and population data

+ Download Google Trends data with the TrendEcon package for individual keywords

+ Aggregate Google Trends data from daily to monthly level 

+ Creating index of multiple keywords

+ Plotting all data

Next steps are creating forecasting model and testing if individual keywords and the composite index make it more accurate
 


### Loading needed libraries

```{r,echo=TRUE, message=FALSE}
library(trendecon)
library(rmarkdown)
library(ggplot2)
library(prophet)
library(lubridate)
library(dplyr)
library(tidyverse)
library(plotly)
library(zoo)
```

### Importing birth and population data

```{r,echo=TRUE, message=FALSE}
birthdata <- read_csv("birthsandpopulationY04Y22.csv",  col_types = cols(timestamp = col_date(format = "%Y-%m-%d")))
ggplot(birthdata, aes(x=timestamp, y=total)) + geom_line() + labs(x="time", y="Number of births", title="Total number of births monthly") 
ggplot(birthdata, aes(x=timestamp, y=first)) + geom_line() + labs(x="time", y="Number of births", title="Number of firstborns monthly") 
```

### Decomposing total births and firstborns
```{r,echo=TRUE, message=FALSE}
births_total<-birthdata$total
births_total <- ts(births_total,start=c(2004,1),end=c(2022,7),frequency=12)
decomp_births_total <- decompose(births_total)
plot(decomp_births_total)

births_firstborn<-birthdata$first
births_firstborn <- ts(births_firstborn,start=c(2004,1),end=c(2022,7),frequency=12)
decomp_births_firstborn <- decompose(births_firstborn)
plot(decomp_births_firstborn)
```


### Seasonally adjusting total births and comparing it to unadjusted data
```{r,echo=TRUE, message=FALSE}

births_total_seasonadj <- births_total - decomp_births_total$seasonal
plot.ts(births_total_seasonadj)
ggplot(birthdata, aes(x=timestamp, y=total)) + geom_line() + labs(x="time", y="Number of births", title="Total number of births monthly") 
```

### Seasonally adjusting firstborns and comparing it to unadjusted data
```{r,echo=TRUE, message=FALSE}
births_firstborn_seasonadj <- births_firstborn - decomp_births_firstborn$seasonal
plot.ts(births_firstborn_seasonadj)
ggplot(birthdata, aes(x=timestamp, y=first)) + geom_line() + labs(x="time", y="Number of births", title="Number of firstborns monthly") 

nt_births_firstborn <- births_firstborn - decomp_births_firstborn$trend
plot(nt_births_firstborn)
```


### Downloading Google trends data for specified keywords and adjusting the seasonality with TrendEcon package

```{r,echo=TRUE, message=FALSE, eval=FALSE}
## Downloading raw data

proc_keyword_init("raskaustesti", "FI")
proc_keyword_init("clearblue", "FI")
proc_keyword_init("ovulaatiotesti","FI")
proc_keyword_init("raskauspahoinvointi","FI")
```

### Creating composite index of the birth keywords
```{r,echo=TRUE, message=FALSE, cache=TRUE}
kw_syntyvyys <- c("raskaustesti","clearblue","ovulaatiotesti","raskauspahoinvointi")

#proc_index(kw_syntyvyys,"FI","syntyvyysindex")
```

### Importing seasonally adjusted data of keywords and turning it to monthly data
```{r,echo=TRUE, message=FALSE}


raskaustesti_OGM <- read_csv("raw/fi/raskaustesti_m.csv", col_types = cols(time = col_date(format = "%Y-%m-%d")))
clearblue_OGM <- read_csv("raw/fi/clearblue_m.csv", col_types = cols(time = col_date(format = "%Y-%m-%d")))
ovulaatiotesti_OGM <- read_csv("raw/fi/ovulaatiotesti_m.csv", col_types = cols(time = col_date(format = "%Y-%m-%d")))
raskauspahoinvointi_OGM <- read_csv("raw/fi/raskauspahoinvointi_m.csv", col_types = cols(time = col_date(format = "%Y-%m-%d")))

```
### Decomposing keywords and removing the trend component of the series
```{r,echo=TRUE, message=FALSE}

sa_kw_raskaustestiOG <- raskaustesti_OGM$value
sa_kw_raskaustestiOG <- ts(sa_kw_raskaustestiOG, start=c(2006,1),end=c(2022,11),frequency=12)
dc_kw_raskaustestiOG <- decompose(sa_kw_raskaustestiOG)
plot(dc_kw_raskaustestiOG)
nt_kw_raskaustesti <- sa_kw_raskaustestiOG - dc_kw_raskaustestiOG$trend
nt_kw_L9_raskaustesti <- lag(nt_kw_raskaustesti,9)
plot(nt_kw_raskaustesti)

sa_kw_clearblueOG <- clearblue_OGM$value
sa_kw_clearblueOG <- ts(sa_kw_clearblueOG, start=c(2006,1),end=c(2022,11),frequency=12)
dc_kw_clearblueOG <- decompose(sa_kw_clearblueOG)
nt_kw_clearblue <- sa_kw_clearblueOG - dc_kw_clearblueOG$trend
plot(dc_kw_clearblueOG)

sa_kw_ovulaatiotestiOG <- ovulaatiotesti_OGM$value
sa_kw_ovulaatiotestiOG <- ts(sa_kw_ovulaatiotestiOG, start=c(2006,1),end=c(2022,11),frequency=12)
dc_kw_ovulaatiotestiOG <- decompose(sa_kw_ovulaatiotestiOG)
nt_kw_ovulaatiotesti <- sa_kw_ovulaatiotestiOG - dc_kw_ovulaatiotestiOG$trend
plot(dc_kw_ovulaatiotestiOG)

sa_kw_raskauspahoinvointiOG <- raskauspahoinvointi_OGM$value
sa_kw_raskauspahoinvointiOG <- ts(sa_kw_raskauspahoinvointiOG, start=c(2006,1),end=c(2022,11),frequency=12)
dc_kw_raskauspahoinvointiOG <- decompose(sa_kw_raskauspahoinvointiOG)
nt_kw_raskauspahoinvointi <- sa_kw_raskauspahoinvointiOG - dc_kw_raskauspahoinvointiOG$trend
plot(dc_kw_raskauspahoinvointiOG)

```



### Plotting the individual seasonally adjusted keywords on monthly level
```{r,echo=TRUE, message=FALSE}
ggplot(raskaustesti_monthly, aes(x=month,y=mean)) + labs(title="Seasonally adjusted keyword index for raskaustesti") + geom_line()
ggplot(clearblue_monthly, aes(x=month,y=mean)) + labs(title="Seasonally adjusted keyword index for clearblue") + geom_line()
ggplot(ovulaatiotesti_monthly, aes(x=month,y=mean)) + labs(title="Seasonally adjusted keyword index for ovulaatiotesti") + geom_line()
ggplot(raskauspahoinvointi_monthly, aes(x=month,y=mean)) + labs(title="Seasonally adjusted keyword index for raskauspahoinvointi") + geom_line()
```

### Creating index for multiple keywords and plotting the index and for comparison the total births index
```{r,echo=TRUE, message=FALSE}
syntyvyysindeksi <- read_csv("raw/fi/syntyvyysindex_sa.csv", col_types = cols(time = col_date(format = "%Y-%m-%d"), 
                                                                              value = col_number()))
syntyvyysindeksi$month <- floor_date(syntyvyysindeksi$time, "month")
syntyvyysindeksi %>% group_by(month) %>% summarize(mean = mean(value))

syntyvyysindeksi_monthly <- (syntyvyysindeksi %>% group_by(month) %>% summarize(mean = mean(value)))

ggplot(syntyvyysindeksi_monthly, aes(x=month,y=mean)) + labs(title="Seasonally adjusted index for birth related keywords") + geom_line()
plot.ts(births_total_seasonadj)

```



### Joining data and adding lags
```{r,echo=TRUE, message=FALSE}
df_nt_births_firstborn <- data.frame(firstborns=as.matrix(nt_births_firstborn),time=as.Date(as.yearmon(time(nt_births_firstborn))))

df_nt_kw_raskaustesti <- data.frame(kw_raskaustesti=as.matrix(nt_kw_raskaustesti),time=as.Date(as.yearmon(time(nt_kw_raskaustesti))))
df_nt_kw_clearblue <- data.frame(kw_clearblue=as.matrix(nt_kw_clearblue),time=as.Date(as.yearmon(time(nt_kw_clearblue))))
df_nt_kw_ovulaatiotesti <- data.frame(kw_ovulaatiotesti=as.matrix(nt_kw_ovulaatiotesti),time=as.Date(as.yearmon(time(nt_kw_ovulaatiotesti))))
df_nt_kw_raskauspahoinvointi <- data.frame(kw_raskauspahoinvointi=as.matrix(nt_kw_raskauspahoinvointi),time=as.Date(as.yearmon(time(nt_kw_raskauspahoinvointi))))

df <- df_nt_births_firstborn %>% left_join(df_nt_kw_raskaustesti, by = c("time"="time"))
df <- df %>% left_join(df_nt_kw_clearblue,by=c("time"="time"))
df <- df %>% left_join(df_nt_kw_ovulaatiotesti,by=c("time"="time"))
df <- df %>% left_join(df_nt_kw_raskauspahoinvointi,by=c("time"="time"))


df %>% plot_ly(., x = ~time) %>% 
  add_trace(y = ~firstborns, name = 'esikoiset',mode = 'lines') %>% 
  add_trace(y = ~kw_raskaustesti, name = 'raskaustesti',mode = 'lines', yaxis="y2") %>% 
  layout(yaxis2=list(overlaying="y", side="right"))

df %>% plot_ly(., x = ~time) %>% 
  add_trace(y = ~firstborns, name = 'esikoiset',mode = 'lines') %>% 
  add_trace(y = ~kw_clearblue, name = 'clearblue',mode = 'lines', yaxis="y2") %>% 
  layout(yaxis2=list(overlaying="y", side="right"))

df %>% plot_ly(., x = ~time) %>% 
  add_trace(y = ~firstborns, name = 'esikoiset',mode = 'lines') %>% 
  add_trace(y = ~kw_ovulaatiotesti, name = 'ovulaatiotesti',mode = 'lines', yaxis="y2") %>% 
  layout(yaxis2=list(overlaying="y", side="right"))

df %>% plot_ly(., x = ~time) %>% 
  add_trace(y = ~firstborns, name = 'esikoiset',mode = 'lines') %>% 
  add_trace(y = ~kw_raskauspahoinvointi, name = 'raskauspahoinvointi',mode = 'lines', yaxis="y2") %>% 
  layout(yaxis2=list(overlaying="y", side="right"))


df2 <- df


df2 <- df2 %>%                           
  mutate(kw_L9_raskaustesti = lag(kw_raskaustesti, n=9,))


df2 %>% plot_ly(., x = ~time) %>% 
  add_trace(y = ~firstborns, name = 'esikoiset',mode = 'lines') %>% 
  add_trace(y = ~kw_L9_raskaustesti, name = 'raskaustesti',mode = 'lines', yaxis="y2") %>% 
  layout(yaxis2=list(overlaying="y", side="right"))



transform(df2,kw_L9_raskaustesti=c(kw_raskaustesti[9],NA))

```

### Selvitä lagin määrä, librarynä forecast aikasarjalla, miten performoi, selittävillä muuttujilla,

