---
title: "Pekka Haimi Master Thesis"
author: "Pekka Haimi"
date: "2022-11-10"
output:
  html_document:
    keep_md: yes
---
This is analysis part of Master Thesis written by Pekka Haimi.  
Aim of the analysis:

* Import Finnish birth and population data

+ Download Google Trends data with the TrendEcon package for individual keywords

+ Aggregate Google Trends data from daily to monthly level 

+ Creating index of multiple keywords

+ Plotting all data

Next steps are creating forecasting model and testing if individual keywords and the composite index make it more accurate
 
library(tsDyn)
library(prophet)
library(lubridate)
library(dplyr)
library(tidyverse)
library(plotly)
library(zoo)
library(tsibble)

library(fpp3)
library(dplyr)
library(tidyverse)
library(tidyquant)
library(stargazer)

library(vars)
library(urca)
library(tsDyn)
library(apt) 
library(timetk)
library(readxl)
library(kableExtra)
library(grid)
library(gridExtra)

### Loading needed libraries

```{r,echo=TRUE, message=FALSE}
library(trendecon)
library(rmarkdown)
library(ggplot2)
library(forecast)
library(gtrendsR)
library(tidyverse)
library(fpp3)
library(kableExtra)
library(vars)

```

### Importing birth and population data and downloading keyword data

```{r,echo=TRUE, message=FALSE}
dataset <- readr::read_csv("birthsandpopulationY04Y22.csv")

          
kw_raskaustesti_raw <- ts_gtrends("raskaustesti",time="2004-01-01 2022-11-30",geo='FI')
kw_raskaustesti_raw <- rename(kw_raskaustesti_raw, raskaustesti = value)

kw_raskausoireet_raw <- ts_gtrends("raskaus oireet",time="2004-01-01 2022-11-30",geo='FI')
kw_raskausoireet_raw <- rename(kw_raskausoireet_raw, raskausoireet = value)

kw_raskaana_raw <- ts_gtrends("raskaana",time="2004-01-01 2022-11-30",geo='FI')
kw_raskaana_raw <- rename(kw_raskaana_raw,raskaana = value)

keywords <- kw_raskaustesti_raw %>% left_join(kw_raskausoireet_raw, by = c("time"="time"))
keywords <- keywords %>% left_join(kw_raskaana_raw, by = c("time"="time"))

dataset <- dataset %>% left_join(keywords,by = c("timestamp"="time"))

dataset <- dataset %>% 
          mutate (Month = yearmonth(timestamp)) %>% 
          dplyr::select(Month,first,total,raskaustesti,raskausoireet,raskaana) %>% 
          as_tsibble(index = Month)

autoplot(dataset)




birthdata <- read_csv("birthsandpopulationY04Y22.csv",  col_types = cols(timestamp = col_date(format = "%Y-%m-%d")))
ggplot(birthdata, aes(x=timestamp, y=total)) + geom_line() + labs(x="time", y="Number of births", title="Total number of births monthly") 
ggplot(birthdata, aes(x=timestamp, y=first)) + geom_line() + labs(x="time", y="Number of births", title="Number of firstborns monthly") 
```

### Order of analysis

*Importing data
*


### Decomposing total births and firstborns
```{r,echo=TRUE, message=FALSE}
births_total<-birthdata$total

##Plotting seasonal changes for total births

birthdata %>% mutate(timestamp=yearmonth(timestamp)) %>%  
  as_tsibble(index=timestamp) -> births_tsibble

births_tsibble %>% gg_season(total,labels="both") +
      labs(y="Total births",
      title ="Seasonal changes of births")

births_tsibble %>%
  gg_subseries(total) +
  labs(
    y = "Total births",
    title = "Seasonal changes of births")


births_total <- ts(births_total,start=c(2004,1),end=c(2022,7),frequency=12)
decomp_births_total <- decompose(births_total)
plot(decomp_births_total)




births_firstborn<-birthdata$first

##Plotting seasonal changes for firstborns


births_tsibble %>% gg_season(first,labels="both") +
      labs(y="Firstborns",
      title ="Seasonal changes of firstborns in Finland")


births_tsibble %>%
  gg_subseries(first) +
  labs(
    y = "Firstborns",
    title = "Seasonal changes of firstborns in Finland")

births_firstborn <- ts(births_firstborn,start=c(2004,1),end=c(2022,7),frequency=12)
decomp_births_firstborn <- decompose(births_firstborn)
plot(decomp_births_firstborn)
```


### Seasonally adjusting total births and comparing it to unadjusted data
```{r,echo=TRUE, message=FALSE}

births_total_seasonadj <- births_total - decomp_births_total$seasonal
plot.ts(births_total_seasonadj)
ggplot(birthdata, aes(x=timestamp, y=total)) + geom_line() + labs(x="time", y="Number of births", title="Total number of births monthly") 
```


### Seasonally adjusting firstborns and comparing it to unadjusted data
```{r,echo=TRUE, message=FALSE}
births_firstborn_seasonadj <- births_firstborn - decomp_births_firstborn$seasonal
plot.ts(births_firstborn_seasonadj)
ggplot(birthdata, aes(x=timestamp, y=first)) + geom_line() + labs(x="time", y="Number of births", title="Number of firstborns monthly") 

nt_births_firstborn <- births_firstborn - decomp_births_firstborn$trend
plot(nt_births_firstborn)
```

### downloading keyword data and applying seasonal adjustment 
```{r,echo=TRUE, message=FALSE}
kw_raskaustesti_raw <- ts_gtrends("raskaustesti",time="2004-01-01 2022-11-30",geo='FI')
kw_raskaustesti_raw <- rename(kw_raskaustesti_raw, raskaustesti = value)

kw_raskaustesti_trend <- tsbox::ts_trend(kw_raskaustesti_raw)
kw_raskaustesti_sa <- tsbox::ts_seas(kw_raskaustesti_raw)
tsbox::ts_plot(kw_raskaustesti_raw)
tsbox::ts_plot(kw_raskaustesti_trend)
tsbox::ts_plot(kw_raskaustesti_sa)

kw_raskausoireet_raw <- ts_gtrends("raskaus oireet",time="2004-01-01 2022-11-30",geo='FI')
kw_raskausoireet_raw <- rename(kw_raskausoireet_raw, raskausoireet = value)
kw_raskausoireet_sa <- tsbox::ts_seas(kw_raskausoireet_raw)

kw_alkuraskaus_raw <- ts_gtrends("alkuraskaus",time="2004-01-01 2022-11-30",geo='FI')
kw_alkuraskaus_raw <- rename(kw_alkuraskaus_raw, alkuraskaus = value)


kw_monitermi_sa_raw <- ts_gtrends_index(c("raskaustesti","raskaus oireet","alkuraskaus"),time="2004-01-01 2022-11-30",geo='FI',sadj=TRUE)
kw_monitermi_sa <- subset(kw_monitermi_sa_raw,id=='sadj')
kw_monitermi_sa <- subset(kw_monitermi_sa,select=-id)
mutate(kw_monitermi_sa, time= as.Date(time, format= "%Y-%m-%d"))
tsbox::ts_plot(kw_monitermi_sa)


tsbox::ts_plot(kw_raskaustesti_nsa)
tsbox::ts_plot(kw_raskaustesti_sa)
```

### Joining data to one table and applying lags




### Joining data and adding lags
```{r,echo=TRUE, message=FALSE}
df_totalbirths <- data.frame(totalbirths=as.matrix(births_total_seasonadj),time=as.Date(as.yearmon(time(births_total_seasonadj))))
df_firstborns <- data.frame(firstborns=as.matrix(births_firstborn_seasonadj),time=as.Date(as.yearmon(time(births_firstborn_seasonadj))))

df <- df_totalbirths %>% left_join(df_firstborns, by = c("time"="time"))
df <- df %>% left_join(kw_raskaustesti_raw, by = c("time"="time"))
df <- df %>% left_join(kw_raskausoireet_raw, by = c("time"="time"))
df <- df %>% left_join(kw_alkuraskaus_raw, by = c("time"="time"))

df <- df %>% relocate(time)
df<- transform(df, raskaustesti = as.numeric(raskaustesti), 
                    raskausoireet = as.numeric(raskausoireet),
               alkuraskaus = as.numeric(raskausoireet))


df %>% plot_ly(., x = ~time) %>% 
  add_trace(y = ~firstborns, name = 'esikoiset',mode = 'lines') %>% 
  add_trace(y = ~raskaustesti, name = 'raskaustesti',mode = 'lines', yaxis="y2") %>% 
  add_trace(y = ~raskausoireet, name = 'raskaus oireet',mode = 'lines', yaxis="y2") %>% 
  add_trace(y = ~alkuraskaus, name = 'alkuraskaus',mode = 'lines', yaxis="y2") %>% 
layout(yaxis2=list(overlaying="y", side="right"))

df %>% mutate(time=yearmonth(time)) %>%  
  as_tsibble(index=time) -> df


df <- df %>%                           
  mutate(raskaustesti_L9 = lag(raskaustesti, n=8,))

df <- df %>%                           
  mutate(raskausoireet_L9 = lag(raskausoireet, n=8,))

df <- df %>%                           
  mutate(alkuraskaus_L9 = lag(alkuraskaus, n=8,))

df <- df %>% filter(time >="2010-01-01")


df %>% plot_ly(., x = ~time) %>% 
  add_trace(y = ~totalbirths, name = 'kaikkisyntyneet',mode = 'lines') %>% 
  add_trace(y = ~raskaustesti_L9, name = 'raskaustesti L9',mode = 'lines', yaxis="y2") %>% 
  add_trace(y = ~raskausoireet_L9, name = 'raskaus oireet L9',mode = 'lines', yaxis="y2") %>% 
  add_trace(y = ~alkuraskaus_L9, name = 'alkuraskaus L9',mode = 'lines', yaxis="y2") %>% 
layout(yaxis2=list(overlaying="y", side="right"))


p1 <- ggplot(df, aes(time, log(totalbirths), color = "Total births")) +
  geom_line() + labs(y = "log of Total births") + theme_bw() + theme(legend.title = element_blank())
p1

df %>%    ggplot(aes(x = time)) +
  geom_line(aes(y = raskaustesti, color = "Raskaustesti")) +
  geom_line(aes(y = raskausoireet, color = "Raskaus Oireet")) +
  geom_line(aes(y = alkuraskaus, color = "Alkuraskaus")) +
  labs(
    title = "Google Trends Queries",
    caption = "Source: Google Trends",
    y = "Search Popularity Index",
    x = "Month"
  )  +
  theme(legend.position = "top", legend.title = element_blank())

```
### Analysing seasonality of the birth data
```{r,echo=TRUE, message=FALSE}

autoplot(df)
### Seasonal changes of keywords
df %>% gg_season(raskaustesti,labels="both") +
      labs(y="Raskaustesti",
      title ="Raskaustesti")

df %>%
  gg_subseries(raskaustesti) +
  labs(
    y = "Raskaustesti",
    title = "Seasonal changes of keyword raskaustesti")

df %>%
  gg_subseries(raskausoireet) +
  labs(
    y = "Raskausoireet",
    title = "Seasonal changes of keyword raskausoireet")


```

### Splitting dataset and building the model
```{r,echo=TRUE, message=FALSE}
#Transforming data frame to tsibble
df <- df %>% as_tsibble(index= time)

train <- df %>% filter(time < '2021-01-01')
test <- df %>% filter(time > '2020-12-01')

fits <- train %>%
  model(
    fit_naive = TSLM(log(totalbirths) ~ 1 + lag(log(totalbirths))),
    fit_GT     = TSLM(
      log(totalbirths) ~ 1 + lag(log(totalbirths)) + log(raskaustesti + 1) + log(raskausoireet + 1) + log(alkuraskaus + 1)
    )
  )

tidy(fits)[, c(1, 2, 3, 4, 6)] %>%
  kable(
    format = "html",
    table.attr = "style='width:75%;' ",
    caption = "Model Estimation Results",
    digits = 3
  ) %>%
  kable_classic_2(full_width = F)


fits_split <- train %>%
  model(
    fit_GT_raskaustesti = TSLM(log(totalbirths) ~ 1 + lag(log(totalbirths)) + log(raskaustesti + 1)),
    fit_GT_raskausoireet = TSLM(log(totalbirths) ~ 1 + lag(log(totalbirths)) + log(raskausoireet + 1)),
    fit_GT_alkuraskaus = TSLM(log(totalbirths) ~ 1 + lag(log(totalbirths))+ log(alkuraskaus + 1)),
  )                   


tidy(fits_split)[, c(1, 2, 3, 4, 6)] %>%
  kable(
    format = "html",
    table.attr = "style='width:75%;' ",
    caption = "Model Estimation Results",
    digits = 3
  ) %>%
  kable_classic_2(full_width = F)

#Forecast Estimation and Results

accuracy(fcs, test)[, c(1, 2, 6, 7)] %>%
  kable(
    format = "html",
    table.attr = "style='width:50%;' ",
    caption = "Model Forecast Results",
    col.names = c("model", "type", "MPE", "MAPE"),
    digits = 3,
  ) %>%
  kable_classic_2(full_width = F)


```

### Selvitä lagin määrä, librarynä forecast aikasarjalla, miten performoi, selittävillä muuttujilla,

