---
title: "Pekka Haimi Master Thesis"
author: "Pekka Haimi"
date: "2022-11-10"
output:
  html_document:
    keep_md: yes
    code_folding: hide
---
This is analysis part of Master Thesis written by Pekka Haimi.  
Aim of the analysis:

* Import Finnish birth and population data

* Download Google Trends data with the TrendEcon package for individual keywords

* Aggregate Google Trends data from daily to monthly level 

* Creating index of multiple keywords

* Plotting all data

* Create forecasting model for births without any external data

* Improve the models with Google Keywords and see if it gets better


### Ladataan tarvittavat kirjastot

```{r,echo=TRUE, message=FALSE}
library(trendecon)
library(fpp3)
library(rmarkdown)
library(plotly)
library(forecast)
library(gtrendsR)
library(tidyverse)
library(kableExtra)
library(vars)
library(GGally)

```

### Importing birth and population data and downloading keyword data

```{r,echo=TRUE, message=FALSE}


dataset <- readr::read_csv("birthsandpopulationY04Y22.csv")
dataset <- dataset %>% dplyr::filter(timestamp >= "2009-01-01")

kw_raskaus_raw <- ts_gtrends("raskaus",time="2009-01-01 2022-12-31",geo='FI')
kw_raskaus_raw <- rename(kw_raskaus_raw, raskaus = value)

kw_raskaustesti_raw <- ts_gtrends("raskaustesti",time="2009-01-01 2022-12-31",geo='FI')
kw_raskaustesti_raw <- rename(kw_raskaustesti_raw, raskaustesti = value)

kw_raskausoireet_raw <- ts_gtrends("raskaus oireet",time="2009-01-01 2022-12-31",geo='FI')
kw_raskausoireet_raw <- rename(kw_raskausoireet_raw, raskausoireet = value)

kw_raskaana_raw <- ts_gtrends("raskaana",time="2009-01-01 2022-12-31",geo='FI')
kw_raskaana_raw <- rename(kw_raskaana_raw,raskaana = value)

keywords <- kw_raskaustesti_raw %>% left_join(kw_raskausoireet_raw, by = c("time"="time"))
keywords <- keywords %>% left_join(kw_raskaana_raw, by = c("time"="time"))
keywords <- keywords %>% left_join(kw_raskaus_raw, by = c("time"="time"))

dataset <- dataset %>% full_join(keywords,by = c("timestamp"="time"),keep=TRUE)

dataset <- dataset %>% mutate (Month = yearmonth(timestamp))


dataset <- dataset  %>% 
          dplyr::select(Month,first,total,raskaus,raskaustesti,raskausoireet,raskaana,timestamp) %>% 
          as_tsibble(index = Month)

## adding lags
dataset <- dataset %>% mutate(raskaus_L12 = lag(raskaus, n=12,))
dataset <- dataset %>% mutate(raskaus_L11 = lag(raskaus, n=11,))
dataset <- dataset %>% mutate(raskaus_L10 = lag(raskaus, n=10,))
dataset <- dataset %>% mutate(raskaus_L9 = lag(raskaus, n=9,))
dataset <- dataset %>% mutate(raskaus_L8 = lag(raskaus, n=8,))
dataset <- dataset %>% mutate(raskaus_L7 = lag(raskaus, n=7,))
dataset <- dataset %>% mutate(raskaus_L6 = lag(raskaus, n=6,))
dataset <- dataset %>% mutate(raskaus_L5 = lag(raskaus, n=5,))
dataset <- dataset %>% mutate(raskaus_L4 = lag(raskaus, n=4,))
dataset <- dataset %>% mutate(raskaus_L3 = lag(raskaus, n=3,))

dataset <- dataset %>% mutate(raskaustesti_L12 = lag(raskaustesti, n=12,))
dataset <- dataset %>% mutate(raskaustesti_L11 = lag(raskaustesti, n=11,))
dataset <- dataset %>% mutate(raskaustesti_L10 = lag(raskaustesti, n=10,))
dataset <- dataset %>% mutate(raskaustesti_L9 = lag(raskaustesti, n=9,))
dataset <- dataset %>% mutate(raskaustesti_L8 = lag(raskaustesti, n=8,))
dataset <- dataset %>% mutate(raskaustesti_L7 = lag(raskaustesti, n=7,))
dataset <- dataset %>% mutate(raskaustesti_L6 = lag(raskaustesti, n=6,))
dataset <- dataset %>% mutate(raskaustesti_L5 = lag(raskaustesti, n=5,))
dataset <- dataset %>% mutate(raskaustesti_L4 = lag(raskaustesti, n=4,))
dataset <- dataset %>% mutate(raskaustesti_L3 = lag(raskaustesti, n=3,))

dataset <- dataset %>% mutate(raskausoireet_L12 = lag(raskausoireet, n=12,))
dataset <- dataset %>% mutate(raskausoireet_L11 = lag(raskausoireet, n=11,))
dataset <- dataset %>% mutate(raskausoireet_L10 = lag(raskausoireet, n=10,))
dataset <- dataset %>% mutate(raskausoireet_L9 = lag(raskausoireet, n=9,))
dataset <- dataset %>% mutate(raskausoireet_L8 = lag(raskausoireet, n=8,))
dataset <- dataset %>% mutate(raskausoireet_L7 = lag(raskausoireet, n=7,))
dataset <- dataset %>% mutate(raskausoireet_L6 = lag(raskausoireet, n=6,))
dataset <- dataset %>% mutate(raskausoireet_L5 = lag(raskausoireet, n=5,))
dataset <- dataset %>% mutate(raskausoireet_L4 = lag(raskausoireet, n=4,))
dataset <- dataset %>% mutate(raskausoireet_L3 = lag(raskausoireet, n=3,))

dataset <- dataset %>% mutate(raskaana_L12 = lag(raskaana, n=12,))
dataset <- dataset %>% mutate(raskaana_L11 = lag(raskaana, n=11,))
dataset <- dataset %>% mutate(raskaana_L10 = lag(raskaana, n=10,))
dataset <- dataset %>% mutate(raskaana_L9 = lag(raskaana, n=9,))
dataset <- dataset %>% mutate(raskaana_L8 = lag(raskaana, n=8,))
dataset <- dataset %>% mutate(raskaana_L7 = lag(raskaana, n=7,))
dataset <- dataset %>% mutate(raskaana_L6 = lag(raskaana, n=6,))
dataset <- dataset %>% mutate(raskaana_L5 = lag(raskaana, n=5,))
dataset <- dataset %>% mutate(raskaana_L4 = lag(raskaana, n=4,))
dataset <- dataset %>% mutate(raskaana_L3 = lag(raskaana, n=3,))

```


```{r,echo=TRUE, message=FALSE}
##Filtering data for 2010 onwards
dataset_2010 <- dataset %>% 
          dplyr::filter(Month >= "2010-01-01")

```

```{r,echo=TRUE, message=FALSE}
##peruskuvaajat
dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% autoplot(vars(first,total))
dataset %>% autoplot(vars(raskaustesti,raskausoireet,raskaana))
```

```{r,echo=TRUE, message=FALSE}
## Plotting seasonal stuff
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% model(stl=STL(total)) %>% components() %>% autoplot()
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% model(stl=STL(first)) %>% components() %>% autoplot()

dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% model(stl=STL(raskaus)) %>% components() %>% autoplot()
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% model(stl=STL(raskaustesti)) %>% components() %>% autoplot()
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% model(stl=STL(raskausoireet)) %>% components() %>% autoplot()
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% model(stl=STL(raskaana)) %>% components() %>% autoplot()

##Subseries
dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% gg_season(total, labels = "both")
dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% gg_season(first, labels = "both")

dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% gg_subseries(total)
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% gg_subseries(first)

dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-12-31"))) %>% gg_subseries(raskaus)
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-12-31"))) %>% gg_subseries(raskaustesti)
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-12-31"))) %>% gg_subseries(raskausoireet)
dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-12-31"))) %>% gg_subseries(raskaana)

```

```{r,echo=TRUE, message=FALSE}
### Päällekkäiset kuvaajat

dataset %>% plot_ly(., x = ~timestamp) %>% 
  add_trace(y = ~total, name = 'kaikki syntyneet',mode = 'lines') %>% 
  add_trace(y = ~raskaus, name = 'raskaus',mode = 'lines', yaxis="y2") %>% 
    add_trace(y = ~raskaus_L9, name = 'raskaus L9',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaus_L8, name = 'raskaus L8',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaus_L7, name = 'raskaus L7',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaus_L6, name = 'raskaus L6',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaus_L5, name = 'raskaus L5',mode = 'lines', yaxis="y2") %>%
  add_trace(y = ~raskaus_L4, name = 'raskaus L4',mode = 'lines', yaxis="y2") %>%
  add_trace(y = ~raskaus_L3, name = 'raskaus L3',mode = 'lines', yaxis="y2") %>%
layout(yaxis2=list(overlaying="y", side="right"))


dataset %>% plot_ly(., x = ~timestamp) %>% 
  add_trace(y = ~total, name = 'kaikki syntyneet',mode = 'lines') %>% 
  add_trace(y = ~raskaustesti, name = 'raskaustesti',mode = 'lines', yaxis="y2") %>% 
    add_trace(y = ~raskaustesti_L9, name = 'raskaustesti L9',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaustesti_L8, name = 'raskaustesti L8',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaustesti_L7, name = 'raskaustesti L7',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaustesti_L6, name = 'raskaustesti L6',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaustesti_L5, name = 'raskaustesti L5',mode = 'lines', yaxis="y2") %>%
  add_trace(y = ~raskaustesti_L4, name = 'raskaustesti L4',mode = 'lines', yaxis="y2") %>%
  add_trace(y = ~raskaustesti_L3, name = 'raskaustesti L3',mode = 'lines', yaxis="y2") %>%
layout(yaxis2=list(overlaying="y", side="right"))

dataset %>% plot_ly(., x = ~timestamp) %>% 
  add_trace(y = ~total, name = 'kaikki syntyneet',mode = 'lines') %>% 
  add_trace(y = ~raskausoireet, name = 'raskausoireet',mode = 'lines', yaxis="y2") %>% 
    add_trace(y = ~raskausoireet_L9, name = 'raskausoireet L9',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskausoireet_L8, name = 'raskausoireet L8',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskausoireet_L7, name = 'raskausoireet L7',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskausoireet_L6, name = 'raskausoireet L6',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskausoireet_L5, name = 'raskausoireet L5',mode = 'lines', yaxis="y2") %>%
  add_trace(y = ~raskausoireet_L4, name = 'raskausoireet L4',mode = 'lines', yaxis="y2") %>%
  add_trace(y = ~raskausoireet_L3, name = 'raskausoireet L3',mode = 'lines', yaxis="y2") %>%
layout(yaxis2=list(overlaying="y", side="right"))



dataset %>% plot_ly(., x = ~timestamp) %>% 
  add_trace(y = ~total, name = 'kaikki syntyneet',mode = 'lines') %>% 
  add_trace(y = ~raskaana, name = 'raskaana',mode = 'lines', yaxis="y2") %>% 
    add_trace(y = ~raskaana_L9, name = 'raskaana L9',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaana_L8, name = 'raskaana L8',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaana_L7, name = 'raskaana L7',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaana_L6, name = 'raskaana L6',mode = 'lines', yaxis="y2") %>%
    add_trace(y = ~raskaana_L5, name = 'raskaana L5',mode = 'lines', yaxis="y2") %>%
  add_trace(y = ~raskaana_L4, name = 'raskaana L4',mode = 'lines', yaxis="y2") %>%
  add_trace(y = ~raskaana_L3, name = 'raskaana L3',mode = 'lines', yaxis="y2") %>%
layout(yaxis2=list(overlaying="y", side="right"))




```


```{r,echo=TRUE, message=FALSE}
### Korrelaatiokuvaajat

dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% ggplot(aes(x=total,y=raskaus)) + geom_point() + geom_smooth(method=lm) + labs(x="Elävänä syntyneet lapset kuukausittain",y="Hakusanan raskaus hakuindeksi")


dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% ggplot(aes(x=total,y=raskaustesti)) + geom_point() + geom_smooth(method=lm) + labs(x="Elävänä syntyneet lapset kuukausittain",y="Hakusanan raskaustesti hakuindeksi")


dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% ggplot(aes(x=total,y=raskaustesti_L7)) + geom_point() + geom_smooth(method=lm) + labs(x="Elävänä syntyneet lapset kuukausittain",y="Hakusanan raskaustesti L7 hakuindeksi")


dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% ggplot(aes(x=total,y=raskausoireet)) + geom_point() + geom_smooth(method=lm) + labs(x="Elävänä syntyneet lapset kuukausittain",y="Hakusanan raskaus oireet hakuindeksi")

dataset %>%  dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% ggplot(aes(x=total,y=raskaana)) + geom_point() + geom_smooth(method=lm) + labs(x="Elävänä syntyneet lapset kuukausittain",y="Hakusanan raskaana oireet hakuindeksi")
  
```

```{r,echo=TRUE, message=FALSE}
##lag plots ## Mutta miksi?
dataset %>% 
           dplyr::filter(timestamp >= "2010-01-01") %>% 
            gg_lag(first,geom="point")
dataset %>% 
           dplyr::filter(timestamp >= "2010-01-01") %>% 
            gg_lag(total,geom="point")
dataset %>% gg_lag(raskaustesti,geom="point")
```


```{r,echo=TRUE, message=FALSE}
#No differencing 
dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
        gg_tsdisplay(log(total),
         plot_type='partial',lag=36) + 
        labs(title="not differenced",y="")

#Seasonal ARIMA of logged total
dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
        gg_tsdisplay(difference(log(total),12),
         plot_type='partial',lag=36) + 
        labs(title="seasonally differenced",y="")
                     
#Double differenced log data
dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
        gg_tsdisplay(difference(log(total),12) %>% difference(),
         plot_type='partial',lag=36) + 
        labs(title="seasonally differenced",y="")
```
#C
According to AutoArima for fit without log is 
model = 101 211

```{r,echo=TRUE, message=FALSE}
##Lets build ARIMA model for the total births

fit <- dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% 
      model( 
       ARIMA101210=ARIMA(log(total) ~ 0 + pdq(1,0,1) + PDQ(2,1,0)),
       ARIMA102210=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,0)),
       ARIMA103210=ARIMA(log(total) ~ 0 + pdq(1,0,3) + PDQ(2,1,0)),
       ARIMA101211=ARIMA(log(total) ~ 0 + pdq(1,0,1) + PDQ(2,1,1)),
       ARIMA102211=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1)),
       ARIMA103211=ARIMA(log(total) ~ 0 + pdq(1,0,3) + PDQ(2,1,1)),
       ARIMA101212=ARIMA(log(total) ~ 0 + pdq(1,0,1) + PDQ(2,1,2)),
       ARIMA102212=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,2)),
       ARIMA103212=ARIMA(log(total) ~ 0 + pdq(1,0,3) + PDQ(2,1,2)),
      auto= ARIMA(log(total),stepwise=FALSE,approx=FALSE)                  
      )


fit %>% pivot_longer(everything(), names_to = "Model name",
                     values_to = "Orders")

glance(fit) %>% arrange(AICc) %>% select(.model:BIC)

fit %>% select(ARIMA101211) %>% gg_tsresiduals(lag=36)

##Ljung box
augment(fit) %>%
  filter(.model == "ARIMA101211") %>%
  features(.innov, ljung_box, lag=24, dof=4)

dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2022-07-01"))) %>% model(ARIMA102211=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1))) %>% forecast() %>% autoplot(dataset)
fit(equation)

```

#Time to build our first Dynamic Regression model with different predictors
```{r,echo=TRUE, message=FALSE}
fit_t_raskaustesti <- dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
      model( 
       no_lag=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti)),
       L3=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L3)),
        L4=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L4)),
        L5=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L5)),
        L6=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L6)),
        L7=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L7)),
        L8=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L8)),
        L9=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L9)),
        L10=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L10)),
        L11=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L11)),
        L12=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L12))
      )

fit_t_raskaus <- dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
      model( 
       no_lag=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus)),
       L3=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L3)),
        L4=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L4)),
        L5=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L5)),
        L6=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L6)),
        L7=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L7)),
        L8=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L8)),
        L9=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L9)),
        L10=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L10)),
        L11=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L11)),
        L12=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L12))
      )

fit_t_raskausoireet <- dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
      model( 
       no_lag=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet)),
       L3=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L3)),
        L4=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L4)),
        L5=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L5)),
        L6=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L6)),
        L7=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L7)),
        L8=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L8)),
        L9=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L9)),
        L10=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L10)),
        L11=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L11)),
        L12=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L12))
      )

fit_t_raskaana <- dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
      model( 
       no_lag=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana)),
       L3=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L3)),
        L4=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L4)),
        L5=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L5)),
        L6=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L6)),
        L7=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L7)),
        L8=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L8)),
        L9=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L9)),
        L10=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L10)),
        L11=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L11)),
        L12=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L12))
      )





glance(fit_t_raskaustesti) %>% arrange(AICc) %>% select(.model:BIC)
glance(fit_t_raskaus) %>% arrange(AICc) %>% select(.model:BIC)
glance(fit_t_raskausoireet) %>% arrange(AICc) %>% select(.model:BIC)
glance(fit_t_raskaana) %>% arrange(AICc) %>% select(.model:BIC)



```
### Creating "the" model
```{r,echo=TRUE, message=FALSE}
fit_t <- dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
      model(
       main_model=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L7) + log(raskaus_L7)+ log(raskausoireet_L7)+log(raskaana_L7)),
       Only_arima=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1)),
       raskaustesti=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L7)),
       raskaus=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaus_L7)),
       raskausoireet=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L7)),
       raskaana=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaana_L7)))

       
glance(fit_t) %>% arrange(AICc) %>% select(.model:BIC)       

fit_t_best <- dataset %>% dplyr::filter(timestamp >= "2010-01-01") %>% 
      model( 
       main_model=ARIMA(total ~ pdq(1,0,1) + PDQ (2,1,1) + raskaustesti_L7 + raskaus_L4 + raskausoireet_L6 + raskaana_L3))

report(fit_t_best)


```
### Training model and forecast
```{r,echo=TRUE, message=FALSE}

fit_t_best_train <- dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2020-01-01"))) %>% 
      model( 
        raskausoireet=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskausoireet_L7)),
        ARIMAlog=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1)),
        main_model=ARIMA(log(total) ~ 0 + pdq(1,0,2) + PDQ(2,1,1) + log(raskaustesti_L7) + log(raskaus_L7)+ log(raskausoireet_L7)+log(raskaana_L7)))
       

fit_t_arimas_train <- dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2020-01-01"))) %>% 
      model( 
      ARIMAlog=ARIMA(log(total)),
       ARIMA010010=ARIMA(log(total) ~ pdq(0,1,0) + PDQ(0,1,0)),
       ARIMA110110=ARIMA(log(total) ~ pdq(1,1,0) + PDQ(1,1,0)),
      ARIMA100110=ARIMA(log(total) ~ pdq(1,0,0) + PDQ(1,1,0)),
      ARIMA111111=ARIMA(log(total) ~ pdq(1,1,1) + PDQ(1,1,1))
       )


fit_t_best_train_2 <- dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2020-07-01"))) %>% 
      model( 
       main_model=ARIMA(total))

future <- dataset %>% dplyr::filter(between(timestamp,as.Date("2020-02-01"),as.Date("2022-12-01")))


fit_t_best_train %>% 
  forecast(future) %>% 
  autoplot(dataset,level=95) + facet_wrap(vars(.model), ncol = 2) + 
  guides(colour="none",fill="none",level="none") +
  geom_label(aes(x=yearmonth("2020 Jan"), y=5000, label = paste0("AICc = ", format(AICc))), 
 data = (glance(fit_t_best_train)))
  
fit_t_best_train %>%  forecast(future) %>% accuracy(dataset) %>% select(.model,RMSE:MAPE)

accuracy(fcs) %>%
select(.model, RMSE:MAPE)                       


fit_t_best_train %>% forecast(future) %>% autoplot(dataset,level=NULL)
fit_t_arimas_train %>% forecast(future) %>% autoplot(dataset,level=NULL)
fit_t_best_train_2 %>% forecast(future) %>% autoplot(dataset)




#mallien vertailu template
fit_t_best_train <- dataset %>% dplyr::filter(between(timestamp,as.Date("2010-01-01"),as.Date("2020-01-01"))) %>% 
      model( 
        ARIMAlog=ARIMA(log(total)),
       ARIMAlog_predictors=ARIMA(log(total) ~ raskaustesti_L7 + raskaus_L4 + raskausoireet_L6 + raskaana_L3),
       ARIMA010010=ARIMA(log(total) ~ pdq(1,1,0) + PDQ(1,1,0)),
       ARIMA110110=ARIMA(log(total) ~ pdq(1,1,0) + PDQ(1,1,0)),
       TSLM=TSLM(total ~ raskaustesti_L7 + raskaus_L4 + raskausoireet_L6 + raskaana_L3),
       ETS=ETS(total)
       )

```
### Forecasting
```{r,echo=TRUE, message=FALSE}

forecast(fit_t_best_train, h=36) %>%
  filter(.model=='auto') %>%
  autoplot(dataset) +
  labs(title = "y",
       y="")

forecast(fit, h=36) %>%
  filter(.model=='arima011011') %>%
  autoplot(dataset) +
  labs(title = "y",
       y="")




```

```
### Linear regression part

```{r,echo=TRUE, message=FALSE}

dataset %>% pivot_longer(c(first,total),names_to="Series") %>% autoplot(value)
dataset %>% pivot_longer(c(raskaustesti,raskausoireet,raskaana),names_to="Series") %>% autoplot(value)

dataset %>% ggplot(aes(x = total,y = raskaustesti))+labs(y="Hakuindeksi: Raskaustesti",x="Syntyneet") + geom_point() + geom_smooth(method="lm",se=FALSE)
dataset %>% ggplot(aes(x = total,y = raskaustesti_L9))+labs(y="Hakuindeksi: Raskaustesti L9",x="Syntyneet") + geom_point() + geom_smooth(method="lm",se=FALSE)
dataset %>% ggplot(aes(x = total,y = raskaustesti_L8))+labs(y="Hakuindeksi: Raskaustesti L8",x="Syntyneet") + geom_point() + geom_smooth(method="lm",se=FALSE)
dataset %>% ggplot(aes(x = total,y = raskaustesti_L7))+labs(y="Hakuindeksi: Raskaustesti L7",x="Syntyneet") + geom_point() + geom_smooth(method="lm",se=FALSE)
dataset %>% ggplot(aes(x = total,y = raskaustesti_L6))+labs(y="Hakuindeksi: Raskaustesti L6",x="Syntyneet") + geom_point() + geom_smooth(method="lm",se=FALSE)


dataset %>% ggplot(aes(x = total,y = raskausoireet))+labs(y="Hakuindeksi: Raskausoireet",x="Syntyneet") + geom_point() + geom_smooth(method="lm",se=FALSE)
dataset %>% ggplot(aes(x = total,y = raskaana))+labs(y="Hakuindeksi: Raskaana",x="Syntyneet") + geom_point() + geom_smooth(method="lm",se=FALSE)
dataset %>% GGally::ggpairs(columns=2:6)


## Linear regression equation (TSLM)
#Naive model
fits <- train %>% 
            model(
              naive=TSLM(log(total) ~ log(lag(total))),
              naive_12 = TSLM(log(total) ~ log(lag(total)) + log(lag(total,12))),
              GT_raskaustesti = TSLM(log(total) ~ log(lag(total)) + log(lag(total,n=12))+log(raskaustesti))
              )

tidy(fits)[, c(1, 2, 3, 4, 6)] %>%
  kable(
    format = "html",
    table.attr = "style='width:75%;' ",
    caption = "Model Estimation Results",
    digits = 3
  ) %>%
  kable_classic_2(full_width = F)
             
fit_fcs <- fits %>% forecast(new_data=test)
fit_fcs %>% autoplot(dplyr::filter(dataset,timestamp> "2010-01-01"),level=NULL)

fc_models <- models %>% forecast(new_data=dataset)
fcs_split %>%
  autoplot(filter(file, year(Month) > 2016), level = NULL)


fc_models %>%
  autoplot()

  fc_models[, c(1, 2, 6, 7)] %>%
  kable(
    format = "html",
    table.attr = "style='width:50%;' ",
    caption = "Model Forecast Results",
    col.names = c("Model", "Type", "MPE", "MAPE"),
    digits = 3
  ) %>%
  kable_classic_2(full_width = F)
  
   
 
 
``` 

```{r,echo=TRUE, message=FALSE}  

###Stupid testing
dataset %>% model(TSLM(total ~ raskaustesti)) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12))) %>% report()
dataset %>% model(TSLM(log(total) ~ log(lag(total)) + log(lag(total,n=12)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=9)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=8)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=7)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=6)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=5)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=4)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=3)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=2)))) %>% report()
dataset %>% model(TSLM(total ~ lag(total) + lag (total,n=12) + log(lag(raskaustesti,n=1)))) %>% report()


dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=12)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=11)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=10)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=9)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=8)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=7)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=6)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=5)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=4)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=3)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=2)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskausoireet,n=1)))) %>% report()


dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=12)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=11)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=10)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=9)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=8)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=7)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=6)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=5)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=4)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=3)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=2)))) %>% report()
dataset %>% model(TSLM(log(total) ~ lag(total) + lag (total,n=12) + log(lag(raskaana,n=1)))) %>% report()

dataset %>% model(TSLM(total ~ raskausoireet)) %>% report()
dataset %>% model(TSLM(total ~ raskausoireet_L9)) %>% report()
dataset %>% model(TSLM(total ~ raskausoireet_L8)) %>% report()
dataset %>% model(TSLM(total ~ raskausoireet_L7)) %>% report()
dataset %>% model(TSLM(total ~ raskausoireet_L6)) %>% report()

dataset %>% model(TSLM(total ~ raskaana)) %>% report()
dataset %>% model(TSLM(total ~ raskaana_L9)) %>% report()
dataset %>% model(TSLM(total ~ raskaana_L8)) %>% report()
dataset %>% model(TSLM(total ~ raskaana_L7)) %>% report()
dataset %>% model(TSLM(total ~ raskaana_L6)) %>% report()

fit_total_keywords <- dataset %>% model(tslm=TSLM(total ~ raskaustesti + raskausoireet + raskaana))
report(fit_total_keywords)

fit_total_keywords_L9 <- dataset %>% model(tslm=TSLM(total ~ raskaustesti_L9 + raskausoireet_L9 + raskaana_L9))
report(fit_total_keywords_L9)

fit_total_keywords_L8 <- dataset %>% model(tslm=TSLM(total ~ raskaustesti_L8 + raskausoireet_L8 + raskaana_L8))
report(fit_total_keywords_L8)

fit_total_keywords_L7 <- dataset %>% model(tslm=TSLM(total ~ raskaustesti_L7 + raskausoireet_L7 + raskaana_L7))
report(fit_total_keywords_L7)

## Fitted values

augment(fit_total_keywords) %>% ggplot(aes(x=Month)) + geom_line(aes(y=total,colour="Data")) + geom_line(aes(y=.fitted,colour="Fitted"))+labs(y=NULL,title="Syntyneet") + scale_colour_manual(values=c(Data="black",Fitted="#D55E00")) + guides(colour=guide_legend(title=NULL))

augment(fit_total_keywords_L9) %>% ggplot(aes(x=Month)) + geom_line(aes(y=total,colour="Data")) + geom_line(aes(y=.fitted,colour="Fitted"))+labs(y=NULL,title="Syntyneet") + scale_colour_manual(values=c(Data="black",Fitted="#D55E00")) + guides(colour=guide_legend(title=NULL))


```
```{r,echo=TRUE, message=FALSE}

### Korrelaatio
Tutkitaan eri viiveiden välistä korrelaatiota syntyvyysaineistoon

```{r,echo=TRUE, message=FALSE}

ggscatterstats(data = dataset, x = total, y = raskausoireet)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L3)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L4)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L5)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L6)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L7)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L8)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L9)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L10)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L11)
ggscatterstats(data = dataset, x = total, y = raskausoireet_L12)


ggscatterstats(data = dataset, x = total, y = raskaus_oireet)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L3)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L4)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L5)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L6)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L7)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L8)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L9)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L10)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L11)
ggscatterstats(data = dataset, x = total, y = raskaus_oireet_L12)

ggscatterstats(data = dataset, x = total, y = raskaustesti)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L3)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L4)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L5)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L6)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L7)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L8)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L9)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L10)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L11)
ggscatterstats(data = dataset, x = total, y = raskaustesti_L12)

ggscatterstats(data = dataset, x = total, y = alkuraskaus)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L3)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L4)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L5)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L6)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L7)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L8)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L9)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L10)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L11)
ggscatterstats(data = dataset, x = total, y = alkuraskaus_L12)
```